---
title: "When are stroke recovery patients most vulnerable to waking or falling asleep?"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    css: styles.css
    social: menu
    source_code: "https://github.com/2zOu2/sleepr-dashboard"
    navbar:
      - { title: "GitHub ↗", href: "https://github.com/2zOu2/sleepr-dashboard", align: right }
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(DT)
library(dplyr)
library(tidyr)
library(htmltools)

# -----------------------------------------------------------------------------
# Load your posterior-summary data here. Expected columns:
#   hour        integer 0-23
#   kind        "sleep" or "wake"
#   mean_dur    predicted episode duration (minutes)
#   lower_95    95% credible interval lower bound
#   upper_95    95% credible interval upper bound
#   n_episodes integer, episodes in hour bin
#
# Example: curves <- readr::read_csv("data/posterior_duration_by_hour.csv")
#
# The dummy below lets the Rmd knit before you wire real data in.
# -----------------------------------------------------------------------------
set.seed(1)
hours <- 0:23
mk <- function(phase, base, amp) {
  rad <- ((hours - phase) / 24) * 2 * pi
  base + amp * cos(rad) + (amp / 6) * cos(rad * 2)
}
curves <- tibble(
  hour       = rep(hours, 2),
  kind       = rep(c("sleep", "wake"), each = 24),
  mean_dur   = c(mk(1, 95, 55), mk(9, 70, 55)),
  lower_95   = c(mk(1, 95, 55) - 22, mk(9, 70, 55) - 22),
  upper_95   = c(mk(1, 95, 55) + 22, mk(9, 70, 55) + 22),
  n_episodes = sample(380:520, 48, replace = TRUE)
)

# Per-participant deviations (for Widget 2 table). Replace with your posterior summary.
participants <- tibble(
  patient_id      = sprintf("P-%03d", c(160,159,56,28,61,43,77,102,115,88,19,204,33,141,7,91)),
  sa_deviation    = round(rnorm(16, 0, 0.25), 2),
  as_deviation    = round(rnorm(16, 0, 0.25), 2),
  sd_across_hours = round(runif(16, 0.20, 0.55), 2),
  n_episodes      = sample(160:310, 16)
)

# Shared plotly config — strip default toolbar clutter.
# Wrap config() calls in this helper instead of splicing a list.
apply_config <- function(p) {
  plotly::config(
    p,
    displaylogo = FALSE,
    modeBarButtonsToRemove = list(
      "pan2d", "select2d", "lasso2d", "zoomIn2d", "zoomOut2d",
      "autoScale2d", "toggleSpikelines", "hoverClosestCartesian",
      "hoverCompareCartesian"
    )
  )
}

# Palette
ACCENT <- "#2a9d8f"  # teal
ACCENT_SOFT <- "rgba(42,157,143,0.18)"
INK    <- "#0b0b0c"
MUTE   <- "#6b7280"
```

Overview {data-icon="fa-chart-line"}
=====================================

Row {data-height=120}
-------------------------------------

### Patients

```{r}
valueBox(50, caption = "Patients (stroke recovery)", icon = "fa-user-group", color = "#f8fafc")
```

### Episodes

```{r}
valueBox("10,715", caption = "Sleep + wake episodes", icon = "fa-moon", color = "#f8fafc")
```

### 95% PI Coverage

```{r}
valueBox("97.3%", caption = "Wake-transition model", icon = "fa-bullseye", color = "#f8fafc")
```

### Median error

```{r}
valueBox("1.6 min", caption = "Sleep-transition model", icon = "fa-ruler", color = "#f8fafc")
```

Row {data-height=480}
-------------------------------------

### Widget 1 — Predicted sleep-episode duration by onset hour (minutes)

```{r widget1}
sleep <- curves %>% filter(kind == "sleep")

p1 <- plot_ly() %>%
  # Night shading 21:00–06:00
  add_trace(
    x = c(21, 24, 24, 21), y = c(0, 0, 200, 200),
    type = "scatter", mode = "none", fill = "toself",
    fillcolor = "rgba(15,23,42,0.05)", hoverinfo = "skip", showlegend = FALSE
  ) %>%
  add_trace(
    x = c(0, 6, 6, 0), y = c(0, 0, 200, 200),
    type = "scatter", mode = "none", fill = "toself",
    fillcolor = "rgba(15,23,42,0.05)", hoverinfo = "skip", showlegend = FALSE
  ) %>%
  # 95% credible interval ribbon
  add_ribbons(
    data = sleep, x = ~hour, ymin = ~lower_95, ymax = ~upper_95,
    fillcolor = ACCENT_SOFT, line = list(color = "transparent"),
    name = "95% credible interval",
    hoverinfo = "skip"
  ) %>%
  # Population line — this is the trace that owns the tooltip
  add_trace(
    data = sleep, x = ~hour, y = ~mean_dur,
    type = "scatter", mode = "lines+markers",
    line = list(color = ACCENT, width = 3),
    marker = list(color = ACCENT, size = 6),
    name = "Population average",
    text = ~sprintf(
      paste0(
        "<b>Sleep episode · %02d:00</b><br>",
        "Predicted duration: <b>%d min</b><br>",
        "95%% CI: [%d, %d] min<br>",
        "Episodes in bin: %s"
      ),
      hour, round(mean_dur), round(lower_95), round(upper_95),
      formatC(n_episodes, big.mark = ",")
    ),
    hoverinfo = "text"
  ) %>%
  layout(
    xaxis = list(
      title = "Hour of sleep onset (24h clock)",
      tickmode = "array", tickvals = seq(0, 24, 3),
      ticktext = sprintf("%02d:00", seq(0, 24, 3)),
      gridcolor = "#eef0f3", zeroline = FALSE
    ),
    yaxis = list(
      title = "Predicted episode duration (minutes)",
      gridcolor = "#eef0f3", zeroline = FALSE, range = c(0, 200)
    ),
    hoverlabel = list(
      bgcolor = "white", bordercolor = ACCENT,
      font = list(family = "ui-monospace, monospace", size = 12, color = INK)
    ),
    margin = list(l = 60, r = 20, t = 10, b = 50),
    paper_bgcolor = "white", plot_bgcolor = "white",
    legend = list(orientation = "h", x = 0, y = 1.08, font = list(size = 11))
  ) %>%
  apply_config()

p1
```

### Widget 1b — Survival curves by start hour

```{r}
# P(episode >= t | starts at hour h) — Weibull approximation for illustration
tgrid <- seq(0, 150, by = 5)
start_hours <- c(1, 5, 11, 17, 21)
pal <- c("#3b82f6", "#0ea5e9", "#eab308", "#f97316", "#a855f7")

surv <- purrr::map2_dfr(
  start_hours, seq_along(start_hours),
  function(h, i) {
    scale <- (curves$mean_dur[curves$kind == "sleep" & curves$hour == h]) * 1.1
    s <- exp(-(pmax(0.01, tgrid) / scale) ^ 1.3)
    tibble(t = tgrid, s = s, start = sprintf("%02d:00", h))
  }
)

p1b <- plot_ly()
for (i in seq_along(start_hours)) {
  h <- start_hours[i]
  d <- surv %>% filter(start == sprintf("%02d:00", h))
  p1b <- p1b %>% add_trace(
    data = d, x = ~t, y = ~s,
    type = "scatter", mode = "lines",
    line = list(color = pal[i], width = 2.25),
    name = sprintf("%02d:00", h),
    hovertemplate = paste0(
      "<b>Start ", sprintf("%02d:00", h), "</b><br>",
      "Duration ≥ %{x} min<br>",
      "Probability: <b>%{y:.0%}</b><extra></extra>"
    )
  )
}
p1b <- p1b %>% layout(
  xaxis = list(title = "Episode duration (minutes)", gridcolor = "#eef0f3", zeroline = FALSE),
  yaxis = list(title = "Survival probability", tickformat = ".0%",
               gridcolor = "#eef0f3", zeroline = FALSE, range = c(0, 1)),
  hoverlabel = list(bgcolor = "white", font = list(family = "ui-monospace, monospace", size = 12)),
  legend = list(title = list(text = "Start hour"), orientation = "v", x = 1.02, y = 1),
  margin = list(l = 60, r = 110, t = 10, b = 50),
  paper_bgcolor = "white", plot_bgcolor = "white"
) %>% apply_config()

p1b
```

Row {data-height=360}
-------------------------------------

### Widget 2 — Per-participant circadian deviations (searchable, sortable)

```{r widget2}
participants %>%
  transmute(
    `Patient ID` = patient_id,
    `Sleep-to-Wake Deviation (log-hazard)` = sa_deviation,
    `Wake-to-Sleep Deviation (log-hazard)` = as_deviation,
    `SD Across Hours` = sd_across_hours,
    `Episodes (n)` = n_episodes
  ) %>%
  datatable(
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 8, dom = "tip",
      columnDefs = list(list(className = "dt-right", targets = 1:4)),
      order = list(list(3, "desc"))
    ),
    class = "compact stripe hover",
    caption = htmltools::tags$caption(
      style = "caption-side: bottom; text-align: left; font-size: 11px; color: #6b7280;",
      "Positive Sleep-to-Wake deviation = individual wakes sooner than the population average at that hour. Click a column header to sort; use the row above each column to filter."
    )
  ) %>%
  formatStyle(
    "Sleep-to-Wake Deviation (log-hazard)",
    background = styleColorBar(c(-0.6, 0.6), "#bfe3de"),
    backgroundSize = "98% 60%", backgroundRepeat = "no-repeat", backgroundPosition = "center"
  ) %>%
  formatStyle(
    "Wake-to-Sleep Deviation (log-hazard)",
    background = styleColorBar(c(-0.6, 0.6), "#fdd0a2"),
    backgroundSize = "98% 60%", backgroundRepeat = "no-repeat", backgroundPosition = "center"
  )
```

Row {data-height=80}
-------------------------------------

### {.no-padding}

<div class="impact">
<b>Why this matters —</b> Quantifying <i>when</i> stroke recovery patients are most vulnerable to fragmented sleep gives clinicians a time-targeted window for intervention — medication scheduling, environmental cues, sleep-hygiene prompts — that a single nightly sleep score cannot. Applied at scale, this shifts rehabilitation from retrospective sleep summaries to prospective, hour-level guidance personalized to each patient's circadian profile.
</div>


About the data {data-icon="fa-database"}
=====================================

Column
-------------------------------------

### Dataset

**Source:** SleepR Study, NIH Award R01NR018979 — source data is confidential and is not redistributed in this repository. **Sample:** 50 adult inpatient and outpatient stroke rehabilitation patients in Atlanta, GA contributed 10,715 contiguous sleep/wake episodes recorded between 2021 and 2024. **Collection:** Wrist actigraphy was sampled continuously at one-minute epochs, classified into binary sleep/wake states using a validated threshold algorithm, and collapsed into episodes pooled across five nights by clock-time of onset. **Variables:** Each episode is described by its onset hour (0–23), posterior mean duration (minutes), 95% credible-interval bounds, episode count, and per-patient log-hazard random effects for the sleep→wake and wake→sleep transitions. **Processing:** Episodes truncated by recording boundaries were right-censored; 24-hour and 12-hour Fourier harmonics of onset hour were pre-computed per episode as fixed-effect covariates.


Model {data-icon="fa-chart-area"}
=====================================

Column
-------------------------------------

### Bayesian hierarchical Weibull hazard

A continuous-time Weibull proportional hazards model for episode-level sleep→wake (SA) and wake→sleep (AS) transitions, with cyclic AR(1) random effects to capture person-specific circadian profiles.

$$h_r(t \mid \eta_n) = \kappa_r\,\exp(\eta_n)\,t^{\kappa_r - 1}, \qquad r \in \{\text{SA}, \text{AS}\}$$

$$\eta_n = \alpha + b_{h_n,\,i_n} + \beta_{c_1} c^{(1)}_n + \beta_{s_1} s^{(1)}_n + \beta_{c_2} c^{(2)}_n + \beta_{s_2} s^{(2)}_n$$

Hierarchical cyclic correlation across hours borrows strength across neighbouring hours and across midnight; weakly-informative priors encode gradual circadian variation and moderate subject-level deviation. Fit in Stan, 4 chains × 2,000 iterations, all R̂ < 1.01.


Impact {data-icon="fa-bullseye"}
=====================================

Column
-------------------------------------

### Why this project matters

Quantifying <i>when</i> stroke recovery patients are most vulnerable to fragmented sleep gives clinicians a time-targeted window for intervention — medication scheduling, environmental cues, sleep-hygiene prompts — that a single nightly sleep score cannot. Applied at scale, this shifts rehabilitation from retrospective sleep summaries to prospective, hour-level guidance personalized to each patient's circadian profile.

### Source code

<p>Full source, data-prep scripts, Stan model code, and this dashboard's Rmd live at
<a href="https://github.com/zhengyiou/sleepr-dashboard" target="_blank">github.com/zhengyiou/sleepr-dashboard ↗</a>.</p>
